TEMPERATURES AND METALLICITIES FOR 10,000 M DWARFS\\ IN THE \apogee\ SURVEY

### 2018-08-13 paper/draft.tex
notes; some restructuring

%%==========================================================================================
\section{Experimental Results} \label{sec:results}

\subsection{Training Sample}

For each model, two tests of consistency were used to assess the validity of our derived test labels: (1) the self-test comparing training labels and test labels, and (2) the leave-one-out cross validation test, in which we train a model on all sources but \emph{n}, then apply the \emph{N-1} source trained model to get the labels for \emph{n}. Precision (scatter) and bias of the model for each test are calculated as the standard deviation and mean of the difference in training and test labels respectively.

@@

\subsubsection{Temperature/Metallicity Model \label{subsec:mann_results}}
For the physical parameter model we trained \thecannon\ on 87 M dwarfs with a two-dimensional temperature/metallicity labels, to a precision of 77K/0.09dex as estimated by the cross validation scatter, similar to the original training sample uncertainties of 60K/0.08dex. We note for this model 6 out of 87 sources show strong visible rotational line broadening (as indicated by the red circles in Figure \ref{fig:mann_validation}), while the remaining sources are mostly old ($>8$ Gyr) and likely slowly rotating sources. While these sources with high rotational broadening have high $\chi^2$ values (those sources with $\chi^2 > 80,000$ in Figure \ref{fig:chi_dist}), we find that the scatter of labels derived for these sources is consistent with the rest of the sample and we do not remove them from the training sample.

@@

\subsubsection{Spectral Type Model \label{subsec:west_results}}
We trained \thecannon\ on 51 M dwarfs with a one-dimensional spectral type label, and obtained a precision of $\pm$0.6 spectral types, more precise than the original training label uncertainty of $\pm$1 spectral type. We note however, that the training sample is distributed heavily towards earlier type sources, with a median spectral type of 3 and only one M8 and one M9 source. As seen in Figure \ref{fig:west_validation}, the model performs well at re-deriving the labels for the M8 and M9 source when they are included in the model (less than 1 SPT difference), but badly when left out (over 2 SPT difference), which indicates \thecannon\ does a poor job at extrapolating to labels outside of the training sample space.

@@

@@

@@

\item[-] Discussion of validation tests and ground truth
\end{enumerate}

>>>>

%%==========================================================================================
\section{Experiments and Results} \label{sec:results}

\subsection{Temperature/Metallicity Model \label{subsec:mann_results}}
For the physical parameter model we trained \thecannon\ on 87 M dwarfs with a two-dimensional temperature/metallicity labels, to a precision of 77K/0.09dex as estimated by the cross validation scatter, similar to the original training sample uncertainties of 60K/0.08dex. We note for this model 6 out of 87 sources show strong visible rotational line broadening (as indicated by the red circles in Figure \ref{fig:mann_validation}), while the remaining sources are mostly old ($>8$ Gyr) and likely slowly rotating sources. While these sources with high rotational broadening have high $\chi^2$ values (those sources with $\chi^2 > 80,000$ in Figure \ref{fig:chi_dist}), we find that the scatter of labels derived for these sources is consistent with the rest of the sample and we do not remove them from the training sample.

Two tests of consistency were used to assess the validity of our derived test labels: (1) the self-test comparing training labels and test labels, and (2) the leave-one-out cross validation test, in which we train a model on all sources but \emph{n}, then apply the \emph{N-1} source trained model to get the labels for \emph{n}. Precision (scatter) and bias of the model for each test are calculated as the standard deviation and mean of the difference in training and test labels respectively. JB: POINT TO THESE RESULTS FOR THIS MODEL IN THE FIGURES

@@

@@

\subsection{Spectral Type Model \label{subsec:west_results}}
We trained \thecannon\ on 51 M dwarfs with a one-dimensional spectral type label, and obtained a precision of $\pm$0.6 spectral types, more precise than the original training label uncertainty of $\pm$1 spectral type. We note however, that the training sample is distributed heavily towards earlier type sources, with a median spectral type of 3 and only one M8 and one M9 source. As seen in Figure \ref{fig:west_validation}, the model performs well at re-deriving the labels for the M8 and M9 source when they are included in the model (less than 1 SPT difference), but badly when left out (over 2 SPT difference), which indicates \thecannon\ does a poor job at extrapolating to labels outside of the training sample space.

JB: NOW POINT HERE TO THE RESULTS OF THE CROSS-VALIDATION AND ETC.

@@

JB: NOW POINT TO RESULTS FIGURES AND SUMMARIZE THEM, IF THERE IS ANYTHING ELSE?

\subsection{Validation of parameters}

@@

JB ALSO POINT TO FIGURES THAT CONNECT TYPES TO PARAMETERS. AND DISCUSS.

JB DO WE WANT A FIGURE THAT SHOWS ABUNDANCES AS A FUNCTION OF MILKY WAY POSITION OR ACTIONS?

@@

\item[-] Discussion of validation tests and ground truth

\item[-] HOGG: Discussion of the implications of these results for the NASA \tess\ Mission and how the future might change..?
\end{enumerate}
